---
title: "Blank_Heatmap"
output: html_document
date: '2022-07-27'
---
---
title: "Teaching_Heatmap"
output: html_document
date: '2022-07-27'
---
---
title: "t_test"
author: "Inhan Lee"
date: "7/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

slides 1,2

Heatmap and Dendrogram: 
A heat map is a graphical representation of data where the individual values contained in a matrix are represented as colors.

Heat maps are typically used in molecular biology to represent the level of expression of many genes across a number of comparable samples (e.g. cells in different states, samples from different patients)

A dendrogram is a tree diagram used to illustrate the arrangement of the clusters produced by hierarchical clustering. 
Dendrograms are often used in computational biology to illustrate the clustering of genes or samples.


## miRcore R camp week 2 day 1

## let's review

### read smalldata
```{r}
dir()
smalldata <- read.table("small_data.txt", header=T, row.names=1, stringsAsFactors=F)
str(smalldata)
```

Goal: Create a heatmap for the 10 Tumor and 10 Control samples in myData

First we need to install an R package to make these pretty plots!
```{r}
install.packages('pheatmap')
```
After installing any package, we must load the library of that package once during your R session
```{r}
library(pheatmap)
```
If you are not able to install pheatmap, you can use heatmap2() instead.
```{r}
?pheatmap
```

For a heatmap, data input needs to be a numeric matrix.
Lets check that the data is a numeric matrix:
```{r}
is.matrix(smalldata)
```
Lets convert our data into a numeric matrix

### select specific rows or columns
### have a matrix form of smalldata
```{r}
smalldata_heatmap <- as.matrix(smalldata)
str(smalldata_heatmap)
is.matrix(smalldata_heatmap)
```
We can log transform the data. Why do we have the +1?
```{r}
smalldata_heatmap = log(smalldata_heatmap+1)
```

Lets make a small smalldata to work with- Select rows with just miRa, miR_c and miR_k
```{r}
m_smalldata_heatmap <- smalldata_heatmap[c(1, 3, 11),]
m_smalldata_heatmap
```

# make pheatmap
Lets use  the pheatmap() function to create a heatmap figure for our small sample dat
```{r}
pheatmap(m_smalldata_heatmap, scale="row", cluster_cols=F, border_color=NA,
         main="Heatmap of Sample Data")

```

Group discussion:

Which two miRs are clustered closely? miR_a and miR_c are clustered together
Is this related to the correlation values? Yes, the correlation value for both is roughly 7.388.

Lets look at the correllation of miR_a and miR_c
define vectors

```{r}
cor(
  m_smalldata_heatmap[1,], m_smalldata_heatmap[2,])
```

# make a scatter plot
```{r}
plot(
  m_smalldata_heatmap[1,], m_smalldata_heatmap[2,],
  xlab="miR_a", ylab="miR_c", col="purple"
)
```

What do you think the correllation will be here? 
Lets check!

```{r}

```
What about miR_a and miR_k?

```{r}
cor(
  m_smalldata_heatmap[1,], m_smalldata_heatmap[3,])
```

Now lets try to cluster both rows and columns: check ?pheatmap() for more information

```{r}
pheatmap(m_smalldata_heatmap, scale="row", border_color=NA, main="Heatmap of Sample Data")

#lets make it look pretty!
pheatmap(smalldata_heatmap, scale="row", cellwidth=15, cellheight=10,
         color=colorRampPalette(c("blue","orange","white"))(256),
         border_color="pink", main="Heatmap of Sample Data")

```

Group time: Everyone share your screen in turn and show your heatmap! 
Do not worry if you are having trouble- your group leaders can help :)

advanced task: add classes of samples (Tumor vs Control) in the heatmap? Search!

```{r}
pheatmap(smalldata_heatmap, scale="row", cellwidth=15, cellheight=10,
         color=colorRampPalette(c("blue","orange","white"))(256),
         border_color="pink", main="Heatmap of Sample Data", )
```

```{r}

```























