---
title: "Iteration and Conditional"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Iterations
Prompt: Let's say I want to type the statement "My favorite number is 1" with 100 different numbers. How do I do this? 

First let's try printing the statement manually: 
```{r}
print("My Favorite Number is 1")
print("My Favorite Number is 2")
print("My Favorite Number is 3")
print("My Favorite Number is 4")
```

I don't know about you, but I'm starting to get a bit bored of copying, pasting and changing the number...let's see if we can utilize the for loop

```{r}
for (i in 1:10) {
  print(i)
}
```

The first part dictates how many times the loop run for. The variable i starts with a value of 1 and will increase it's value after it finishes the loop.

i is like a temporary variable; it disappears after the for loop is done running

Let's apply this to our printing problem: 
```{r}
for (i in 1:10) {
  print( paste("My Favorite Number is", i, sep="... ") )
}
```

### Conditional
What if I only like even numbers? How can I reprogram the earlier program so only even numbers print? 
```{r}
for (i in 1:10) {
  if ( i%%2 == 0 ) {print(i)}
  else {print("Odd Corruption... I refuse to print!")}
}
```

Example #2, let's make a vector that creates a vector of headers that go like GroupA_1, GroupB_1, GroupA_2, GroupB_2...GroupA_10, GroupB_10
```{r}
headers <- vector()
for (i in 1:10) {
  headers <-append(headers,
    paste("GroupA", i, sep="_")
  )
  headers <- append(headers,
    paste("GroupB", i, sep="_")
  )
}
print(headers)
```

Let's make a vector containing the pvalues from the ttests of smalldata tumors vs. controls: 
```{r}

```







Your turn: 
Go to your Group!
Using small data
1. Change the smalldata to a matrix.
```{r}
#smalldata <- read.table("small_data.txt", ...)
smalldata <- as.matrix(smalldata)
```

2. Make an empty data.frame named ttest.
```{r}
ttest <- data.frame()
ttest
```

3. For the first gene (hint: first row) 
      1. calculate t.test p-value between expression values all tumor samples (hint: first 10 samples and the 
         values need to be numeric) and all control samples (change from your earlier task)
```{r}
tumor1 <- smalldata[1, 1:10]
control1 <- smalldata[1, 11:20]
firstP <- t.test(tumor1, control1)$p.value
firstP
```
         
      2. If p-value is less than 0.0001, add to the expression values in the ttest (hintt: use rbind 
         function).
4. Do the step 3 for the remaining genes (total 19) using a for loop
```{r}
pvalues <- c(firstP)
for (i in 2:20) {
  tumor_i <- smalldata[i, 1:10]
  control_i <- smalldata[i, 11:20]
  pval_i <- t.test(tumor_i, control_i)$p.value
  if (pval_i < 0.0001) {
    #ttest <- rbind(ttest, pval_i)
  }
  pvalues <- append(pvalues, pval_i)
}
ttest

###other way
#pval = vector()
#for(i in 1:nrow(smalldataM))
#{
#  pval[i] = t.test(smalldataM[i,grep("Control", colnames(smalldataM))],smalldataM[i, grep("Tumor", #colnames(smalldataM))])$p.value
#}
#pval
ttest$expression <- pvalues  #this will not work if ttest is barely initialized
ttest
```

5. Add a column for the ratio between means for the tumors and control (tumor_mean/control_mean) (to the ttest dataframe)
```{r}
ratio <- vector()
for (i in 1:nrow(smalldata)) {
  tumors_i <- smalldata[i, grep("Tumor", colnames(smalldata))]
  control_i <- smalldata[i, grep("Control", colnames(smalldata))]
  #ratio <- append(ratio, mean(tumors_i) / mean(control_i) )
  ttest <- rbind(ttest, mean(tumors_i)/mean(control_i))
}
colnames(ttest) <- c("ratio")
head(ttest)
```

6. Create a new data frame called updata. Add the data for all the miRs where the tumor mean is HIGHER than the control mean among miR with p<0.0001.
```{r}
updata <- data.frame()
for (i in 1:nrow(ttest)) {
  if (ttest[i,]$ratio > 1) {
    updata <- rbind(updata, ttest[i,] )
  }
}
updata
```













